Measuring Directionality in Double-Beta Decay and Neutrino Interactions with Kiloton-Scale Scintillation Detectors

### 2014-04-21 RefereeResponse/RefereeResponse_v4.1.tex

\documentclass[11pt]{article}

\usepackage[usenames,dvips]{pstcol}

\usepackage{color}

\usepackage{amssymb}

\usepackage{amsmath}

\usepackage{graphicx,color}

%

%\input{symbols.tex}

%

\newcounter{question}

\newcommand{\noteq}{{\vskip0.1in\bf \large \color{red} [Q:{\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\noteQ}{{\bf \LARGE \color{blue} [Q: {\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\notea}{{\bf \large \color{red} [A]~}}

\newcommand{\noteA}{{\bf \LARGE \color{blue} [A]~}}

\newcommand{\noteqq}[1]{{\bf \large \color{red} [Q: {\color{black}#1}]~}}

\newcommand{\noteaa}[1]{{\bf \large \color{red} [A: {\color{black}#1}]~}}

%

\textwidth 6.5in

\textheight 9.4in

\topmargin -0.8in

\oddsidemargin 0.0in

\evensidemargin 0.0in

\parindent 0.0in

\def\baselinestretch{1.0}

\def \Russ {Translated from Russian. Fixed. Thanks.}

\def \F    {Fixed.}

\def \FT   {Fixed. Thanks.}

\def \D    {Done.}

\def \AP   {Basically Stylistic- Author's Prerogative.}

\begin{document}

\begin{flushright}

\today

\end{flushright}

%----------------------------------------------------------------

Dear JINST Editors,\\

\parindent 1.0in

We thank the referees for their very helpful and knowledgeable
comments. We have substantially rewritten the manuscript in response 
to the comments as detailed in the list below.

There are several areas in which the referees requested
substantially expanding the scope of the paper.  We envisioned this
paper as a proof of  the principle that the use of the fast timing
photodetectors presently being developed would enable the separation
of Cherenkov and scintillation light, as the reconstruction of the
direction of low energy ($\sim$1 MeV) electrons in a liquid
scintillator detector has never been shown before.  In this paper, we
want to: 1) show scintillation/Cherenkov separation is possible; 2)
determine the parameters of a large scintillator detector that are
important to this separation; and 3) demonstrate that reconstruction
techniques based on Cherenkov light would work in idealized
situations.

We are now working on two follow-up papers that are substantially
beyond the scope of the present paper, moving beyond
proof-of-principle into the laborious and time-intensive area of
reconstruction algorithms, areas for which the difficulty is known
well by some of us from our experience on KamLand. The first is a
detailed study of reconstruction in more realistic scenarios. The
second is a proposal for a $0\nu\beta\beta$ experiment  where the detailed
background reduction factors and sensitivity to new physics are
presented. SuperNEMO has a similar paper, but for scintillator
detectors nothing similar exists since a directional signal has not
been considered for them in the past.  Finally, we will be having a
turnover of young personnel so this paper provides an accessible summary
of their work.

\parindent 3.0in

    Sincerely,\\

\parindent 4.0in

		         Lindley Winslow, for the Authors

\parindent 0.0in

%\begin{verbatim}

%\end{verbatim}

\noteq [Referee 1 Intro] ``I like  very much the clean, transparent, simple
and well-motivated methodology (with one exception..). The structure
of the paper is nice''

\notea{\bf Many thanks.}\\

\noteq [Referee 1 Intro] ``However the presentation of the results is a
different thing... I am particularly concerned with the fact...,
instead of showing the resulting CosTheta distribution that would seem
like the obvious result one wants to see. .... Why not to disclose
this openly to allow the reader to evaluate the results in detail?''

\notea{\bf  This is our fault- the desired plot was (is) there
(now Figure 5), but was labeled as $p_x/|\vec{p}|$ rather than 
$\cos\theta$. We have made this clear on the figure and in the
text. We have also added the distribution for the 2.1~MeV point. Thanks.}\\

\noteq [Referee 1: Page 1] ``Please specify FWHM or sigma and give a
reference for both cases.''

\notea{\bf  \D  ~We have added references to KamLAND and Borexino.}\\

\noteq [Referee 1: Page 1] ``...but please avoid misplacing the adjective good in
this context. ...''

\notea{\bf  \FT}\\

\noteq [Referee 1: Page 2] ``..the authors should..check and refer to
the work of NEXT and Super-Nemo collaborations:''

\notea{\bf  \D We have added a discussion of previous work on directionality.}\\

\noteq [Referee 1: Page 2] ``Define PPO here, not later''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3] ``Please define all magnitudes in eq. 2.1 and 2.2''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3]  ``Indicating the direction of the
electrons should not harm''

\notea{\bf  We were not sure what the referee was requesting. If this is a request for figure 1, we have tried adding the arrows and it becomes too busy. We think the caption description works well.}\\

\noteq [Referee 1: Page 3] ``None of the best limits or claims in bb0
uses an isotope that satisfies this condition..''

\notea{\bf \FT We have rewritten this section to address this comment.}\\

\noteq [Referee 1: Page 4] ``could you perhaps include the
attenuation @400nm and n@400..?''

\notea{\bf \D We have added the attenuation length at 400nm. We have also added the n at other wavelengths in other locations. Unfortunately we cannot cite one figure in the references as the index of refraction is taken from the KamLAND simulation which is based on several measurements that are in the cited references.}\\

\noteq [Referee 1: Page 4] ``Could you please include also the cutoff
wavelength previously referred to?''

\notea{\bf This has been added.}\\

\noteq [Referee 1: Page 4] `` Scattering have not been included..''

\notea{\bf We have added a figure and accompanying text to show that
the scattering only adds a late tail. Thanks.}\\

\noteq [Referee 1: Page 4] ``For which digitized values come from ??''

\notea{\bf \F}\\

\noteq [Referee 1: Page 4] ``5 MeV electrons.. It is very disturbing
to benchmark the simulation with electrons having an energy doubling
the Qbb of the most common isotopes..''

\notea{\bf We have clarified in several places 
that the 5 MeV case is for neutrino-electron scattering experiments. For $0\nu\beta\beta$ we
have made it explicit that we consider 1.4 and 2.1 MeV.}\\

\noteq [Referee 1: Page 4] ``Please specify also the track length in
the direction of emission''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``Before the solutes in liquid... Please rephrase''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``I hope that [43] is not the final version
but a draft.

\notea{\bf It was a draft; we have changed the reference. Thanks!}\\

\noteq [Referee 1: Page 6] ``So I would be very much convinced ..if
the authors would report Fig. 3 during the electron history
(equivalently, energy or distance, including several histograms for
each case. It would be nice to see at which energy scale or length the
electron isotropizes and the Cherenkov cone smears out.''

\notea{\bf  In the updated figure 4 (angular PE hit distribution after time cut), we added the histograms for 1.4~MeV and 2.1~MeV. These show that also for 2.1 and 1.4~MeV the PE hits are not `isotropized`.%, including the full history of the electron tracks. 
We have also added the path lengths and total distance and time from the vertex to the text. We also present in Fig.~\ref{truthscatt} of this report  more raw data for the referees which support the statement that the scattering angle is small while the majority of the Cherenkov light is produced.}\\

\begin{figure}
        \begin{center}
        \includegraphics[scale=0.50]{Figure1_1.pdf} 
        \includegraphics[scale=0.50]{Figure1_2.pdf} 
	   \caption[]{For electrons with an initial energy of 1.4~MeV, we show (Top) the angle between the initial electron direction and the direction at 0.7 MeV (=half of the initial electron energy) and (Bottom) the angle between the initial electron direction and the direction of the last Geant4 step before the electron stops. We note that due to scattering the final direction of the electron before it stops does not correspond to the initial direction; however the scattering angle is small while the majority of Cherenkov light is produced.}
	  \label{truthscatt}
        \end{center}
\end{figure}

\noteq [Referee 1: Pages 7-8] ``This is interesting, but specially
section 6 seems a bit off-topic to me. I am not against keeping it...''

\notea{\bf We'd like to keep it. }\\

\noteq [Referee 1: Page 9] ``Please add a histogram of the reconstructed
angle for all energies...''

\notea{\bf  \D}\\

\noteq [Referee 1: Page 9] ``..the vertex reconstruction has to be
obtained from all photons and the direction from the Cherenkov photons..''

\notea{\bf We agree with the referee and plan to address this in future papers. To do a credible and useful job on the position dependence will require a fairly-detailed vertex reconstruction. This is beyond the scope of this present 'proof-of-principle' paper. We have tried to clarify this in the text }\\

\noteq [Referee 1: Fig. 7] ``the authors should show the spread of the
cos(theta) distribution for each of the energies.

\notea{\bf \D\ See Figure 5.}\\

\noteq [Referee 1: Fig. 7 ``It would be nice to show/indicate..also the
Cherenkov threshold for high wave-length (n about 1.4)

\notea{\bf We have added this to the text.}\\

\noteq [Referee 1: Conclusions] ``Should be rewritten after the
previous objections are addressed.''

\notea{\bf \D\ Thanks!}\\

\noteq [Referee 1: Page 7-8] ``I see no convincing proof of it with
the information here presented. Hope the answers help clarifying the issue.''

\notea{\bf We have made the sentence more specific. Thanks.}\\

%\newpage

\noteq [Referee 2: General Comments] ``The paper introduces an
original technical development which could increase the interest of
large scale scintillation detectors for neutrino physics in general
and neutrinoless double beta decay in particular''

\notea{\bf  Thank you.}\\ %, this warms our hearts.}\\

\noteq [Referee 2: General comments] `` ..however the discussion
should deserve...a deeper analysis.''

\notea{\bf Yes. We have included more data from the lower energy studies,
have extracted much more detail from the simulation, and have
extensively rewritten the manuscript. It is still at the
proof-of-principle level. Our next work will include a full sensitivity analysis including a deeper understanding of the performance of the reconstruction algorithms with a more detailed simulation. This is a big job and we are working on it! }\\

\noteq [Referee 2: General Comments] ``The physics case is included in
the title and mentioned in the text but never discussed in some
detail...A description of why the method is powerful for double beta
decay, and..to what extent it can improve the experimental sensitivity
would complete the discussion

\notea{\bf  \D}\\

\noteq [Referee 2: General Comments] ``The chosen example (5 MeV,
head-to-head single electrons) is of little interest for double beta decay.''

\notea{\bf We have added plots and numbers for 1.4 and 2.1 MeV, and
substantially rewritten the text to differentiate between the cases
of neutrino-electron scattering and double-beta decay.}\\

\noteq [Referee 2: General Comments] ``The dependence on the decay
position.. is not discussed.''

\notea{\bf To do a credible and useful job on the position dependence
will require a fairly-detailed vertex reconstruction. This is beyond
the scope of this present `proof-of-principle' paper. We have tried
to clarify this in the text. }\\

\noteq [Referee 2: General Comments] Threshold effects as well as the
effect of statistical fluctuations for the low energy tails.. are not
discussed. The same holds for position and direction reconstruction
capabilities''

\notea{\bf Statistical fluctuations are addressed by the studies at lower energies. We mention explicitly in the conclusions the widening of the distributions due to decreased statistics. Threshold effects are outside of the scope of this paper. In general in these detectors, the largest loss of photon statistics is at the photocathode. The electronics for PMT readout are very efficient. }\\

\noteq [Referee 2: Abstract] ``While appropriate for neutrino physics,
energy resolution is definitely worse with respect to other techniques..''

\notea{\bf We have fixed the text. Thanks.}\\

\noteq [Referee 2: Introduction] ``A reference ... would help to understand
the connection with Majorana neutrinos.''

\notea{\bf \D}\\

\noteq [Referee 2: Page 1] Is this 1 sigma or FWHM?

\notea{\bf It's sigma- we have fixed the text.}\\

\noteq [Referee 2: Page 2] ``better than water Cherenkov detectors''

\notea{\bf \FT}\\

\noteq [Referee 2: Page 2 ] ``strong suppressant of backgrounds''..this
statement should deserve a deeper quantitative discussion ..''

\notea{\bf We agree. However, a useful ``quantitative discussion''
will require a vertex reconstruction. We will get there, but at present
it's beyond the scope of `proof-of-principle, looking to see if we
can see the signal. We have tried to make this clearer in the text. }\\

\noteq [Referee 2: Equ. 2.1] ``..the meaning of the parameter `x' is
not described.''

\notea{\bf \F}\\

\noteq [Referee 2: Page 3] What is the typical cutoff value and the
corresponding useful fraction of Cherenkov light?

\notea{\bf We have added the cutoff values.}\\

\noteq [Referee 2: Page 3] `` `Probable case being...' Events with
asymmetric electron energies are far more abundant.''

\notea{\bf \FT}\\

\noteq [Referee 2: Page 4] `` `Fully absorbing' ''. How are results
influenced by deviations from this hypothesis?''

\notea{\bf This is a level of detail well beyond what we've done so
far; once we have a real detector design and a full simulation it
can be addressed, but at this point devoting effort to it 
doesn't seem productive.}\\

\noteq [Referee 2: Page 4] ``Single 5 MeV electrons''.

\notea{\bf \F}\\

\noteq [Referee 2: Page 4] ``I believe the `realistic' analysis
mentioned here is indispensable ..for this process''

\notea{\bf We agree, and are working on it. It isn't easy, or,
unfortunately, fast compared to a post-doc tenure.}\\

\noteq [Referee 2: Page 6] ``The dependence on the position and the
corresponding transit time variations cannot in my opinion be neglected.''

\notea{\bf This part of the paper describes a simplified test simulation (where the photo-detector TTS was switched off) to extract the time scales of various effects. We made changes in the text to make this clear.}\\

\noteq [Referee 2: Figure 3] `` This figure is very interesting
to understand the capacity of the method to maintain directionality
information... The comparison with a single event (or equivalently the
distribution of the variances) and the distribution for 1 MeV
electrons would improve to understand also the limits/power of the method''

\notea{\bf  We have added to what is now figure 4 the distributions for the two lower energies. We think this is a nice illustration of the limits/power of the method.}\\

\noteq [Referee 2: reconstruction] ``It is not clear if limiting the
reconstruction to Cherenkov photons improve reconstruction''

\notea{\bf We agree. In this paper we were focussing on the Cherenkov photons as a simple way to apply direction reconstruction algorithms. In the future we will work to reconstruct position and direction simultaneously from the scintillation and Cherenkov light.}\\

\end{document}

\noteq [Referee 1]

\notea{\bf }\\
